---
title: "Analysis"
output: html_document
date: "2023-03-15"
---

# __CLEANING DATA__
# get columns names from 3 files
# 2015 - deleting columns (based on their names)
## 2015 - deleting colums we do not need (based on their index)
## 2015 - count columns numbers so we can compare
## 2016 - deleting colums we do not need (based on their index)
## 2016 count columns numbers so we can compare
## 2017 - deleting colums we do not need (based on their index)
## 2017 count columns numbers so we can compare
#obtaining unique names of "country" columns (2016 and 2017, as 2015 doesn't have country data)
# Changed various names for US, UK and Canada to "United States", "United Kingdom" and "Canada" in 2016
# Changed various names for US, UK and Canada to "United States", "United Kingdom" and "Canada" in 2017
# Named countries other than UK,USA and Canada as OTHER (2016 and 2017)
#introducing column "age" in all three datasets (2015, 2016, 2017) and populating it with mutate
#running Janitor for 2015 and 2016
# Renaming candy names column in 2017
## 2015 and 2016: creating a "year" column with mutate and extracting data for it from a timestamp column
# 2017 creating a column "year" and populating it with "2017"
#2016 and 2017 - renaming "gender"columns.
#finding names of columns unique for 2015 and 2016
#unifying candy names in 2015 and 2016
#finding names of columns unique for 2016 and 2017
#unifying candy names in 2015 and 2016
#COMBINING ALL 3 DATASETS with bind_rows, creating a dataset named combine_candy
#comparing number of columns in 3 datasets and a number of columns in combine_candy dataset
#deleting columns we don't need from combine_candy
# Unifying column "are you trick or treating" (because we forgot to do it before; as a result, we united three columns and cleaned values in them from NA)
#checking number of columns in a final dataset
#writing dataset to a file

# __ANALYSIS__

# load libraries
```{r}

library(tidyverse)
library(ggplot2)
library(scales)
```

# read clean data
```{r}

candy_analysis <- read_csv("clean_data/clean_candy.csv")

```

##1. What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}

# What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

candy_analysis_ratings <- candy_analysis %>%
    
    select(-c("Year","age","are_you_trick_or_treating","gender",
              "country", "internal_id"))

   sum(!is.na(candy_analysis_ratings))


```

##2. What was the average age of people who are going out trick or treating?
##3. What was the average age of people who are not going trick or treating?

```{r}

av_age_trick_or_treating <- candy_analysis %>%
group_by(are_you_trick_or_treating) %>%
    filter(age <= 100) %>%
         summarise(average_age = mean(age,na.rm= TRUE))

av_age_trick_or_treating

##average age of people going trick or treating: 35 (rounded)
##average age of people not going trick or treating: 39 (rounded)
```

##5. How many people rated Starburst as despair?
```{r}

filter(candy_analysis, starburst == "DESPAIR") %>% count()


#The number of people that really don't like Starburst (rated it as despair) = 1990
```
#4. For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}

    candy_totals <- candy_analysis_ratings %>%
          gather("candy_bar","ratings") %>%
            group_by(candy_bar, ratings) %>% count(ratings, name = "total") %>%
            arrange(desc(`total`))
    
#run the code below for totals for each rating
head(filter(candy_totals, ratings == "DESPAIR")
head(filter(candy_totals, ratings == "JOY")
head(filter(candy_totals, ratings == "MEH")

```

#For the next three questions, count despair as -1, joy as +1, and meh as 0.
##What was the most popular candy bar by this rating system for each gender in the data set ?
```{r}
candy_favourite_gender <- candy_analysis %>%
    select("Year", "gender", "country")

candy_by_gender <- bind_cols(candy_favourite_gender, candy_analysis_ratings)

##candy_2 will give the most popular candy bar by the rating system for each gender in the data set. 
#men, women and NA all prefer any_sized_candy_bar

candy_2 <- pivot_longer(candy_by_gender, cols=c(4:93),
             names_to = "candy",
             values_to="rating",
             values_drop_na = TRUE) %>%
mutate(rating = case_when(
    rating == "JOY" ~ 1,
    rating == "DESPAIR" ~ -1,
    rating == "MEH" ~ 0)) %>% 
  group_by(gender, candy) %>% 
  summarise(sum=sum(rating)) %>% 
  slice_max(sum, n=3)
```

##What was the most popular candy bar in each year?
```{r}

candy_3 <- pivot_longer(candy_by_gender, cols=c(4:93),
             names_to = "candy",
             values_to="rating",
             values_drop_na = TRUE) %>%
mutate(rating = case_when(
    rating == "JOY" ~ 1,
    rating == "DESPAIR" ~ -1,
    rating == "MEH" ~ 0)) %>% 
  group_by(Year, candy) %>% 
  summarise(sum=sum(rating)) %>% 
  slice_max(sum, n=3)

##candy_3 will give the answer without plotting a graph. It is any_size_candy_bar for all years

```

##What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
```{r}

candy_4 <- pivot_longer(candy_by_gender, cols=c(4:93),
             names_to = "candy",
             values_to="rating",
             values_drop_na = TRUE) %>%
mutate(rating = case_when(
    rating == "JOY" ~ 1,
    rating == "DESPAIR" ~ -1,
    rating == "MEH" ~ 0)) %>% 
  group_by(country, candy) %>% 
  summarise(sum=sum(rating)) %>% 
  slice_max(sum, n=3)

##Candy_4 will give the the answer without plotting a graph
#UK is the only outlier 

```
##What are the top three candy per gender
```{r}


ggplot(data = candy_2, aes(fill=candy, x = gender, y = sum)) +
  geom_col(position="dodge", stat="identity") +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))
```

## Top three candy bars by year
```{r}
ggplot(data = candy_3, aes(fill=candy, x = Year, y = sum)) +
  geom_col(position="dodge", stat="identity") +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))
```
## Top three candy bars by country

```{r}

ggplot(data = candy_4, aes(fill=candy, x = country, y = sum)) +
  geom_col(position="dodge", stat="identity") +
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))

##warning messages appear above but they still run

```

##Proportion of ratings (per year)
```{r}



candy_analysis_year <- candy_analysis %>%
    select(-c("age","are_you_trick_or_treating","gender",
              "country", 'internal_id')) %>%
pivot_longer(cols=c(2:91),
             names_to = "candy",
             values_to="rating",
             values_drop_na = TRUE) %>%
  group_by(Year, rating) %>%
  count()

ggplot(candy_analysis_year, aes(fill=rating, x=Year, y=n)) +
    geom_bar(position="fill", stat="identity") +
labs(title = "Proportion of ratings (per year) ", x = "Year", y = "proportion")
```

##Proportion of ratings (per gender)
```{r}

candy_analysis_gender <- candy_analysis %>%
    select(-c("age","are_you_trick_or_treating","Year",
              "country", 'internal_id')) %>%
pivot_longer(cols=-'gender',
             names_to = "candy",
             values_to="rating",
             values_drop_na = TRUE) %>%
  group_by(gender, rating) %>%
  count()
ggplot(candy_analysis_gender, aes(fill=rating, y=n, x=gender)) +
    geom_bar(position="fill", stat='identity') +
  labs(title = "Proportion of ratings (per gender) ", x = "gender", y = "proportion")
```


##Proportion of ratings (per year)
```{r}
##Proportion of ratings (per year)

candy_analysis_missing_year <- candy_analysis %>%
    select(-c("age","are_you_trick_or_treating","gender",
              "country", 'internal_id')) %>% 
pivot_longer(cols=c(2:91),
             names_to = "candy",
             values_to="rating") %>%
group_by(Year, rating=is.na(rating)) %>% 
  count()

ggplot(candy_analysis_missing_year, aes(fill=rating, x=Year, y=n)) + 
    geom_bar(position="fill", stat="identity") +
labs(title = "Proportion of ratings (per year) ", x = "Year", y = "proportion",
    subtitle = "False = Present data, True = Missing data",
    fill = "Missing or not"
  ) 

## Proportion of missing values per gender
candy_analysis_missing_gender <- candy_analysis %>%
    select(-c("age","are_you_trick_or_treating","Year",
              "country", 'internal_id')) %>% 
pivot_longer(cols=-'gender',
             names_to = "candy",
             values_to="rating") %>%
group_by(gender, rating = is.na(rating)) %>% count() 




```

##Proportion of missing values per year
```{r}

```{r}

candy_analysis_missing_year <- candy_analysis %>%
    select(-c("age","are_you_trick_or_treating","gender",
              "country", 'internal_id')) %>% 
pivot_longer(cols=c(2:91),
             names_to = "candy",
             values_to="rating") %>%
group_by(Year, rating=is.na(rating)) %>% 
  count()
ggplot(candy_analysis_missing_year, aes(fill=rating, x=Year, y=n)) + 
    geom_bar(position="fill", stat="identity") +
labs(title = "Proportion of ratings (per year) ", x = "Year", y = "proportion", fill = "Missing or not", subtitle = "False = Present data, True = Missing data")+
  scale_y_continuous(label = label_percent())
```

```{r}
```
##Proportion of missing values per country
```{r}
candy_analysis_missing_country <- candy_analysis %>%
    select(-c("age","are_you_trick_or_treating","Year",
              "gender", 'internal_id')) %>% 
pivot_longer(cols=-'country',
             names_to = "candy",
             values_to="rating") %>%
group_by(country, rating=is.na(rating)) %>% 
  count()
ggplot(candy_analysis_missing_country, aes(fill=rating, x=country, y=n)) + 
    geom_bar(position="fill", stat="identity") +
labs(title = "Proportion of ratings (per country)", x = "country", y = "proportion", fill = "Missing or not", subtitle = "False = Present data, True = Missing data")+
  scale_y_continuous(label = label_percent())

```
##Proportion of missing values per gender
```{r}
candy_analysis_missing_gender <- candy_analysis %>%
    select(-c("age","are_you_trick_or_treating","Year",
              "country", 'internal_id')) %>% 
pivot_longer(cols=-'gender',
             names_to = "candy",
             values_to="rating") %>%
group_by(gender, rating=is.na(rating)) %>% 
  count()
# plot
ggplot(candy_analysis_missing_gender, aes(fill=rating, x=gender, y=n)) + 
    geom_bar(position="fill", stat="identity") +
labs(title = "Proportion of ratings (per gender)", x = "gender", y = "proportion", fill = "Missing or not", subtitle = "False = Present data, True = Missing data")+
  scale_y_continuous(label = label_percent())
```
